---
title: "Boulder Chickadee Study Data Exploration"
author: "Kathryn Grabenstein"
date: "12/1/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Study Goals

The Boulder Chickadee Study is a long term study exploring the natural history of black-capped and mountain chickadees in Boudler, CO. The study was established in 2019 with wooden nest boxes mounted throughout Boulder County, CO to encourarge chickadees to breed in the boxes. Boxes help researchers monitor breeding and access nestlings.

The goal of this document is to begin preliminary analysis of breeding. Specifically, to create maps of nest boxes, heat maps of nest box occupancy, and other reproductive output information, such as, timing of breeding, reproductive success, etc. 

```{r, include=FALSE}
#import libraries
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(smatr)
library(lme4)
library(devtools)
library(dplyr)
library(lubridate)
library(raster)
library(sf)


#set working directory
setwd("~/Documents/Data Analysis/Chapter5")

```


```{r, include = FALSE}
#import data 
#import ALL nest boxes
nest.boxes <- read.csv("ChickadeeNestBoxes_ANT - ChickadeeNestBoxes.csv")

#import csv for each years 
chickadee.nests <- read.csv("ChickadeeNests_2019-2021.csv")

str(chickadee.nests)

#reformat dates to be continuous variable (not string) for clutch initiation, hatch and Day 12
{chickadee.nests$Hatch.Date <- as.Date(chickadee.nests$Hatch.Date, format = "%d-%b-%y")
chickadee.nests$Clutch.Initiation <- as.Date(chickadee.nests$Clutch.Initiation, format = "%d-%b-%y")
chickadee.nests$Day12 <- as.Date(chickadee.nests$Day12, format = "%d-%b-%y")
}

#calculate julian date to make plotting make sense (MAKE IT MAKE SENSE PLEASE)
chickadee.nests$Julian <- yday(chickadee.nests$Hatch.Date) 

#calculate julian date to make plotting make sense (MAKE IT MAKE SENSE PLEASE)
chickadee.nests$Julian.CI <- yday(chickadee.nests$Clutch.Initiation) 



class(chickadee.nests$Hatch.Date)
class(chickadee.nests$Clutch.Initiation)
class(chickadee.nests$Day12)

unique(chickadee.nests$Species)
unique(chickadee.nests$Site)
unique(chickadee.nests$Year)

#need to get lat/longs from nest box datasheet
chickadee.nests <- left_join(chickadee.nests, nest.boxes, by = c('Nest.Box.ID'))
chickadee.nests

chickadee.nests$Elevation <- as.numeric(chickadee.nests$Elevation)
chickadee.nests$Year <- factor(chickadee.nests$Year)

#make new variable: nestlings/clutch (proportion eggs)
chickadee.nests$Prop.Nestlings <- chickadee.nests$Nestlings/chickadee.nests$Clutch.Size

str(chickadee.nests)


```



## Boulder Chickadee Study Maps
The Boulder Chickadee Study is located in Boulder County, CO. Maps show current nest box locations, as well as where black-capped and mountain chickadees have nested over the past three years.


```{r, include= FALSE}
#register key bc google 
register_google(key = "AIzaSyCmSK6DOJIIqs2IrbzYQTZaUF9J33GXTFA")

Longitude <- chickadee.nests$Longitude
Latitude <- chickadee.nests$Latitude

Longitude.2 <- nest.boxes$Longitude
Latitude.2 <- nest.boxes$Latitude

#get map
chickadeemap <- get_map(location = c(lon = -105.4, lat = 40), 
                        maptype= "satellite", 
                        source ="google", zoom = 10, 
                        color = 'bw' )
#get map and make rectangular

chickadeemap.rectangle <- ggmap(get_map(location = c( -105.35,  40.05), 
                                zoom=10, scale="auto", color = 'bw', maptype= "satellite", 
                                source ="google" )) +
                                scale_x_continuous(limits = c(-105.0,-105.7), expand = c(0, 0)) +
                                scale_y_continuous(limits = c(39.9,40.2), expand = c(0, 0))
chickadeemap.rectangle


  

heat.map.2 <-  get_map(location = c(lon = -105.30, lat = 40), 
                        maptype= "satellite", 
                        source ="google", zoom = 12, 
                        color = 'bw' )

heat.map.3 <-  get_map(location = c(lon = -105.53, lat = 40.03), 
                        maptype= "satellite", 
                        source ="google", zoom = 12, 
                        color = 'bw' )

```

## Map of all nest boxes (n = 373)
```{r maps, echo = FALSE}
#Map of ALL nest boxes in Boulder Chickadee Study


ggmap(chickadeemap)

#plot points
{mapPoints<- ggmap(chickadeemap) + 
  geom_point(aes(x = Longitude.2, y = Latitude.2), 
             data = nest.boxes, size = 4, shape = 21, alpha = 0.8, fill = "royalblue1") 
mapPoints
}

```

```{r, elevation cross section, make the raster} 

library(raster)
library(rgdal)
library(ggplot2)
library(sp)
install.packages("topoDistance")
library(topoDistance)

raster <- raster("USGS_13_n41w106_20220331.tif")
plot(raster)

raster.2 <- raster("USGS_13_n40w106_20220216.tif")
plot(raster.2)

merged <- merge(raster, raster.2)
plot(merged)

box <- as(extent(-105.7, -105.1, 39.8, 40.15), 'SpatialPolygons')
cropped.raster <- crop(x = merged, y = box)
plot(cropped.raster)

#check coordinate reference system
raster@crs

#format points for raster plotting
nest.boxes$Latitude <- as.numeric(nest.boxes$Latitude)
nest.boxes$Longitude  <- as.numeric(nest.boxes$Longitude)

nest.boxes <- na.omit(nest.boxes) 


# plot banding locations
{points <- cbind(nest.boxes$Longitude , nest.boxes$Latitude )
nestbox.points = SpatialPoints(points, proj4string=CRS("+init=epsg:4326"))
pts3.nestboxes = spTransform(nestbox.points, projection(raster))
head(coordinates(pts3.nestboxes))
}

#make line for transect
to.draw <- SpatialPoints(rbind(c(-105.6, 40.05), c(-105.2, 40)))
sp_line <- as(to.draw,"SpatialLines")

myCol <- grey.colors(6)

{plot(cropped.raster, col = myCol)
plot(pts3.nestboxes, add=TRUE)
plot(sp_line, add = TRUE)
}

topoProfile(
  cropped.raster,
  sp_line,
  pts = 100,
  type = "base",
  singlePlot = FALSE,
  rows = NULL,
  cols = NULL,
  limits = NULL,
  legendx = "bottomright"
)



```
```{r, plot elevational differences between MOCH and BCCH}

chickadee.nests %>%
  ggplot(aes(Elevation.x, fill=Species)) +
  geom_density(alpha=0.2) +
   theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), axis.line = element_line(colour = "black")) 

```





## Maps of chickadee nests

```{r,echo=FALSE}
#plot nests only, color by species to create heat map

#Map of all nests
{map.nests <- ggmap(chickadeemap) +
    geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                      data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue"))
map.nests
}

#heat map of points
{map.nests <- ggmap(chickadeemap) + 
              geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                            data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
              stat_density2d(aes(x = Longitude, y = Latitude, color = Species), 
                             alpha = 0.8,
                             data = chickadee.nests) +
              scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
              scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) 
            
map.nests
}


#heat map of points
{map.nests.2 <- ggmap(chickadeemap) + 
              geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                            data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
              geom_density2d(aes(x = Longitude, y = Latitude, color = Species), 
                             alpha = 0.4,
                             data = chickadee.nests) +
              scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
              scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
              facet_wrap(vars(Year))
            
map.nests.2
}


```

## Summary Statistics for Breeding Phenology & Reproductive Success


```{r}
#summary statistics
#Number of nests per species
chickadee.nests %>% 
  group_by(Species) %>% 
  tally()

#number of nests per year
chickadee.nests %>% 
  group_by(Year) %>% 
  tally()

#number of nests per species per year
chickadee.nests %>% 
  group_by(Year, Species) %>% 
  tally()


```




## Breeding Success

```{r}
#What is average clutch size per species?
chickadee.nests %>%
  group_by(Species) %>%
  summarize(Mean = mean(Clutch.Size, na.rm=TRUE), Median = median(Clutch.Size, na.rm=TRUE),
            sd = sd(Clutch.Size, na.rm=TRUE))

##Does clutch size differ between years?
chickadee.nests %>%
  group_by(Year, Species) %>%
  summarize(Mean = mean(Clutch.Size, na.rm=TRUE), Median = median(Clutch.Size, na.rm=TRUE),
            sd = sd(Clutch.Size, na.rm=TRUE))

```


## Breeding Phenology Plots

### Does clutch size differ between black-capped and mountain chickadees? 
```{r,echo=FALSE}
#Does clutch size differ by species?

#clutch size by species
{violin.chickadee <- ggplot(chickadee.nests, aes(x= Species, y=Clutch.Size, fill = Species)) + 
    geom_violin() 
  
  
  #add summary statistics
  violin.chickadee + 
    geom_boxplot(position=position_dodge(0.9), width=0.2) + 
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    scale_x_discrete(labels = c('Black-capped','Mountain')) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), axis.line = element_line(colour = "black"))
}

#Does clutch size differ by species?
#strip chart
#clutch size by species
{strip.chickadee <- ggplot(chickadee.nests, aes(x= Species, y=Clutch.Size, fill = Species)) + 
    #geom_boxplot(width=0.3,lwd=1.5) + 
    #geom_violin(trim = FALSE) +
    geom_boxplot(notch = F, alpha = 0)+
    geom_jitter(position=position_jitter(0.1), size=5, shape = 21, alpha = 0.6) 
  
  
  #add summary statistics
  strip.chickadee + 
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_classic() +
    scale_x_discrete(labels = c('Black-capped','Mountain')) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), axis.line = element_line(colour = "black"))
}

#Does clutch size differ by species?
#strip chart
#clutch size by species
{strip.chickadee.2 <- ggplot(chickadee.nests, aes(x= Species, y=Clutch.Size, fill = Species)) + 
    #geom_boxplot(width=0.3,lwd=1.5) + 
    #geom_violin(trim = FALSE) +
    geom_boxplot(notch = F, alpha = 0)+
    geom_jitter(position=position_jitter(0.1), size=5, shape = 21, alpha = 0.6) 
  
  
  #add summary statistics
  strip.chickadee.2 + 
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_classic() +
    scale_x_discrete(labels = c('Black-capped','Mountain')) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), axis.line = element_line(colour = "black")) +
    facet_wrap(vars(Year))
}



```

### Does clutch size differ by elevation, grouped by species?

```{r,echo= FALSE}

#Does clutch size differ by elevation, grouped by species?
#Scatterplot 
#clutch size by elevation 
{scatter.chickadee.2 <- ggplot(chickadee.nests, aes(x=Elevation, y=Clutch.Size, fill = Species)) +
    geom_jitter(size=5, alpha = 0.6, shape = 21, height = 0) +
    geom_smooth(method = lm, aes(color=Species), fill = "grey") +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_bw() 
    #theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          #panel.grid.minor = element_blank(), axis.title.x=element_blank(), 
         # axis.line = element_line(colour = "black")) +
    
    
scatter.chickadee.2
}

{scatter.chickadee <- ggplot(chickadee.nests, aes(x=Elevation, y=Clutch.Size, fill = Species)) +
    geom_jitter(size=5, alpha = 0.6, shape = 21, height =0) +
    geom_smooth(method = lm, aes(color=Species), fill = "grey") +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_bw() + 
    #theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          #panel.grid.minor = element_blank(), axis.title.x=element_blank(), 
         # axis.line = element_line(colour = "black")) +
    facet_grid(vars(Year))
    
scatter.chickadee
}

```


### Does the proportion of nestlings at Day 12 differ by elevation?

```{r, echo = FALSE}
#Does the number of proportion of nestlings (to account for diff in number of eggs)
#differ by elevation?

#Scatterplot 
#nestlings by elevation 
{scatter.chickadee.nestling <- ggplot(chickadee.nests, aes(x=Elevation, y=Prop.Nestlings, fill = Species)) +
    geom_point(size=5, alpha = 0.6, shape = 21) +
    geom_smooth(method = lm, aes(color=Species), fill = "grey") +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), 
          axis.line = element_line(colour = "black")) 
  
  scatter.chickadee.nestling
}


#Scatterplot 
#nestlings by elevation by year
{scatter.chickadee.nestling.2 <- ggplot(chickadee.nests, aes(x=Elevation, y=Prop.Nestlings, fill = Species)) +
    geom_point(size=5, alpha = 0.6, shape = 21) +
    geom_smooth(method = lm, aes(color=Species), fill = "grey") +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
    theme_bw() + 
    facet_grid(vars(Year))
  
  scatter.chickadee.nestling.2
}


```

### What is average hatch date for species?
```{r, echo = FALSE }
data_subset <- chickadee.nests[ , c("Julian")]   
hatch_date_success <- chickadee.nests[complete.cases(data_subset), ]


hatch_date_count <- hatch_date_success %>% 
  group_by(Species,Year) %>%
  count(Julian)


#need  to make year Factor, not numeric
hatch_date_count$Year <- factor(hatch_date_count$Year)


ggplot(hatch_date_count, aes(x = Julian, y = n, group = Species, fill = Species, color = Species)) +
 geom_point(size=5, alpha = 0.6, shape = 21) +
  #geom_smooth() +
  facet_grid(vars(Species)) +
  scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1))


ggplot(hatch_date_count, aes(x = Julian, y = n, group = Year, color = Year)) +
 geom_point(aes(shape = Year), size=5, alpha = 0.6) +
  geom_line(aes(group = Year)) +
  facet_grid(vars(Species)) +
  #scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  #scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1)) 
  
    
```


### What is average clutch initiation date for species?
```{r}
data_subset_2 <- chickadee.nests[ , c("Julian.CI")]   
CI_date_success <- chickadee.nests[complete.cases(data_subset), ]


CI_date_count <- CI_date_success %>% 
  group_by(Species,Year) %>%
  count(Julian.CI)

ggplot(CI_date_count, aes(x = Julian.CI, y = n, group = Species, fill = Species, color = Species)) +
 geom_point(size=5, alpha = 0.6, shape = 21) +
  geom_smooth() +
  facet_grid(vars(Species)) +
  scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1))


```









### What is average hatch date for each species by year?
```{r, echo = FALSE}

ggplot(hatch_date_count, aes(x = Julian, y = n, fill = Species,  group = Species)) +
 geom_point(size=5, alpha = 0.6, shape = 21) +
  geom_line() +
  facet_grid( Year ~ Species) +
  scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1)) 

```


### How does hatch date vary by elevation?

```{r, echo = FALSE}

{scatter.chickadee.hatch <- ggplot(chickadee.nests, aes(x=Elevation, y=Julian, fill = Species, group = Species)) +
    geom_point(size=5, alpha = 0.6, shape = 21) +
    geom_smooth(method = lm, aes(color=Species), fill = "grey") + 
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1))  

  scatter.chickadee.hatch
}



{scatter.chickadee.hatch <- ggplot(chickadee.nests, aes(x=Elevation, y= Julian, fill = Species, group = Year)) +
    geom_point(size=5, alpha = 0.6, shape = 21) +
    geom_smooth(method = "lm", aes(linetype=Year, color = Species), fill = "grey") + 
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1))  +
  facet_grid(vars(Species))
    
  scatter.chickadee.hatch
}



```


```{r}
### Goal is to plot tarsus v HI 
## What is the relationship between tarsus length and Hybrid index (unscaled)? 

## Import HI for every bird
snps <- read.csv("wgs_gghybrid_loci_1.csv")

#import sequenced adults
adults.sequenced <- read.csv("chickadee_wgs_ADULTS.csv")

#import sequenced nestlings
nestlings.sequenced <- read.csv("Nestlings_to_Exclude.csv")

#import banded adults
adults.banded <- read_csv("BirdsBanded.csv")

#import banded nestlings
nestlings.banded <- read.csv("NestlingsBanded.csv")

str(snps)
str(adults.sequenced)
str(adults.banded)

adults.banded$`Tarsus.(mm)` <- as.numeric(adults.banded$`Tarsus.(mm)`)

#rename adult sequenced column to join
names(adults.sequenced)[1] <- 'USGS.Number'
str(adults.sequenced)

adults.sequenced$USGS.Number <- as.character(adults.sequenced$USGS.Number)

adults.morpho <- left_join(adults.sequenced, adults.banded, by = c('USGS.Number'))
  
names(snps)[1] <- 'USGS.Number'
snps$USGS.Number <- as.character(snps$USGS.Number)

adults.morpho.HI <- left_join(adults.morpho, snps,  by = c('USGS.Number'))

adults.morpho.HI$`Tarsus.(mm)` <- as.numeric(adults.morpho.HI$`Tarsus.(mm)`)
adults.morpho.HI$hi.h.posterior.mode <- as.numeric(adults.morpho.HI$hi.h.posterior.mode)


#Scatterplot 
#tarsus by HI
{scatter.tarsus.hi <- ggplot(adults.morpho.HI, aes(x=`Tarsus.(mm)`, y=hi.h.posterior.mode, fill = Species)) +
    geom_point(size=5, alpha = 0.6, shape = 21) +
    #geom_boxplot(notch = F, alpha = 0) +
    scale_fill_manual(values = c("rosybrown2","cornflowerblue","cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "cornflowerblue","cornflowerblue")) +
    geom_vline(xintercept = 17.88) +
    geom_vline(xintercept = 19.81) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), 
          axis.line = element_line(colour = "black")) 
  
  scatter.tarsus.hi
}


{scatter.tarsus <- ggplot(adults.banded, aes(x=Species, y=`Tarsus.(mm)`, fill = Species)) +
    geom_jitter(size=5, alpha = 0.6, shape = 21) +
    geom_boxplot(notch = F, alpha = 0) +
    scale_fill_manual(values = c("rosybrown2","purple","cornflowerblue")) +
    scale_color_manual(values = c("rosybrown2", "purple","cornflowerblue")) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), 
          axis.line = element_line(colour = "black")) 
  
  scatter.tarsus
}


```



```{r}
adults.banded$Species <- as.factor(adults.banded$Species)


adults.banded %>% 
  dplyr::group_by(Species) %>%
  dplyr::summarize(mean_tarsus = mean(`Tarsus.(mm)`, na.rm = TRUE))


```




