---
title: "Boulder Chickadee Study Data Exploration"
author: "Kathryn Grabenstein"
date: "12/1/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Study Goals

The Boulder Chickadee Study is a long term study exploring the natural history of black-capped and mountain chickadees in Boudler, CO. The study was established in 2019 with wooden nest boxes mounted throughout Boulder County, CO to encourarge chickadees to breed in the boxes. Boxes help researchers monitor breeding and access nestlings.

The goal of this document is to begin preliminary analysis of breeding. Specifically, to create maps of nest boxes, heat maps of nest box occupancy, and other reproductive output information, such as, timing of breeding, reproductive success, etc. 

```{r, include=FALSE}
#import libraries
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(smatr)
library(lme4)
library(devtools)
library(dplyr)
library(lubridate)
library(raster)
library(sf)


#set working directory
setwd("~/Documents/Data Analysis/Chapter5")

```


```{r, include = FALSE}
#import data 
#import ALL nest boxes
nest.boxes <- read.csv("ChickadeeNestBoxes_ANT - ChickadeeNestBoxes.csv")

#import csv for each years 
chickadee.nests <- read.csv("ChickadeeNests_2019-2021.csv")

str(chickadee.nests)

#reformat dates to be continuous variable (not string) for clutch initiation, hatch and Day 12
{chickadee.nests$Hatch.Date <- as.Date(chickadee.nests$Hatch.Date, format = "%d-%b-%y")
chickadee.nests$Clutch.Initiation <- as.Date(chickadee.nests$Clutch.Initiation, format = "%d-%b-%y")
chickadee.nests$Day12 <- as.Date(chickadee.nests$Day12, format = "%d-%b-%y")
}

#calculate julian date to make plotting make sense (MAKE IT MAKE SENSE PLEASE)
chickadee.nests$Julian <- yday(chickadee.nests$Hatch.Date) 

#calculate julian date to make plotting make sense (MAKE IT MAKE SENSE PLEASE)
chickadee.nests$Julian.CI <- yday(chickadee.nests$Clutch.Initiation) 


#need to get lat/longs from nest box datasheet
chickadee.nests <- left_join(chickadee.nests, nest.boxes, by = c('Nest.Box.ID'))

#make new variable: nestlings/clutch (proportion eggs)
chickadee.nests$Prop.Nestlings <- chickadee.nests$Nestlings/chickadee.nests$Clutch.Size


```

## Boulder Chickadee Study Maps
The Boulder Chickadee Study is located in Boulder County, CO. Maps show current nest box locations, as well as where black-capped and mountain chickadees have nested over the past three years.


```{r, include= FALSE}


Longitude <- chickadee.nests$Longitude
Latitude <- chickadee.nests$Latitude



#get map
chickadeemap <- get_map(location = c(lon = -105.4, lat = 40), 
                        maptype= "satellite", 
                        source ="google", zoom = 10, 
                        color = 'bw' )
#get map and make rectangular


```

## Map of all nest boxes (n = 373)
```{r maps, echo = FALSE}
#Map of ALL nest boxes in Boulder Chickadee Study


ggmap(chickadeemap)

#plot points
{mapPoints<- ggmap(chickadeemap) + 
  geom_point(aes(x = Longitude, y = Latitude), 
             data = nest.boxes, size = 4, shape = 21, alpha = 0.8, fill = "royalblue1") 
mapPoints
}

```

```{r, elevation cross section, make the raster} 

library(raster)
library(rgdal)
library(ggplot2)
library(sp)
install.packages("topoDistance")
library(topoDistance)

raster <- raster("USGS_13_n41w106_20220331.tif")
plot(raster)

raster.2 <- raster("USGS_13_n40w106_20220216.tif")
plot(raster.2)

merged <- merge(raster, raster.2)
plot(merged)

box <- as(extent(-105.7, -105.1, 39.8, 40.15), 'SpatialPolygons')
cropped.raster <- crop(x = merged, y = box)
plot(cropped.raster)

#check coordinate reference system
raster@crs

#format points for raster plotting
nest.boxes$Latitude <- as.numeric(nest.boxes$Latitude)
nest.boxes$Longitude  <- as.numeric(nest.boxes$Longitude)

nest.boxes <- na.omit(nest.boxes) 


# plot banding locations
{points <- cbind(nest.boxes$Longitude , nest.boxes$Latitude )
nestbox.points = SpatialPoints(points, proj4string=CRS("+init=epsg:4326"))
pts3.nestboxes = spTransform(nestbox.points, projection(raster))
head(coordinates(pts3.nestboxes))
}

#make line for transect
to.draw <- SpatialPoints(rbind(c(-105.6, 40.05), c(-105.2, 40)))
sp_line <- as(to.draw,"SpatialLines")

myCol <- grey.colors(6)

{plot(cropped.raster, col = myCol)
plot(pts3.nestboxes, add=TRUE)
plot(sp_line, add = TRUE)
}

topoProfile(
  cropped.raster,
  sp_line,
  pts = 100,
  type = "base",
  singlePlot = FALSE,
  rows = NULL,
  cols = NULL,
  limits = NULL,
  legendx = "bottomright"
)



```
```{r, plot elevational differences between MOCH and BCCH}

chickadee.nests %>%
  ggplot(aes(Elevation.x, fill=Species)) +
  geom_density(alpha=0.2) +
   theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.title.x=element_blank(), axis.line = element_line(colour = "black")) 

```





## Maps of chickadee nests

```{r,echo=FALSE}
#plot nests only, color by species to create heat map

#Map of all nests
{map.nests <- ggmap(chickadeemap) +
    geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                      data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
    scale_fill_manual(values = c("rosybrown2", "cornflowerblue"))
map.nests
}

#heat map of points
{map.nests <- ggmap(chickadeemap) + 
              geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                            data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
              stat_density2d(aes(x = Longitude, y = Latitude, color = Species), 
                             alpha = 0.8,
                             data = chickadee.nests) +
              scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
              scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) 
            
map.nests
}


#heat map of points
{map.nests.2 <- ggmap(chickadeemap) + 
              geom_point(aes(x = Longitude, y = Latitude, fill = Species), 
                            data = chickadee.nests, size = 5, alpha = 0.6, shape = 21) +
              geom_density2d(aes(x = Longitude, y = Latitude, color = Species), 
                             alpha = 0.4,
                             data = chickadee.nests) +
              scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
              scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
              facet_wrap(vars(Year))
            
map.nests.2
}


```

## Summary Statistics for Breeding Phenology & Reproductive Success


```{r}
#summary statistics
#Number of nests per species
chickadee.nests %>% 
  group_by(Species) %>% 
  tally()

#number of nests per year
chickadee.nests %>% 
  group_by(Year) %>% 
  tally()

#number of nests per species per year
chickadee.nests %>% 
  group_by(Year, Species) %>% 
  tally()


```



## Breeding Phenology Plots


### What is average clutch initiation date for species?
```{r}
data_subset_2 <- chickadee.nests[ , c("Julian.CI")]   
CI_date_success <- chickadee.nests[complete.cases(data_subset), ]


CI_date_count <- CI_date_success %>% 
  group_by(Species,Year) %>%
  count(Julian.CI)

ggplot(CI_date_count, aes(x = Julian.CI, y = n, group = Species, fill = Species, color = Species)) +
 geom_point(size=5, alpha = 0.6, shape = 21) +
  geom_smooth() +
  facet_grid(vars(Species)) +
  scale_fill_manual(values = c("rosybrown2", "cornflowerblue")) +
  scale_color_manual(values = c("rosybrown2", "cornflowerblue")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 315, vjust = 0.5, hjust=1))

